---
title: "xyloplot examples"
author: "Daniel Greene"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Examples}
  %\VignetteEngine{knitr::rmarkdown}
  %\usepackage[utf8]{inputenc}
---
```{r echo=FALSE}
knitr::opts_chunk$set(dev="svg", fig.width=7, fig.height=5)
library(xyloplot)
knit_print.xyloplot <- function(x, ...) xyloplot:::plot.xyloplot(x)
set.seed(1)
```

The `xyloplot` function displays continuous or discrete data as histogram style violin plots, or "xylophones".

`xyloplot` accepts a simple numeric vector...
```{r}
xyloplot(rnorm(1000))
```

... or a list of simple numeric vectors...
```{r}
xyloplot(
 x=lapply(1:3, function(mean) rnorm(mean=mean, n=1000)), 
 breaks=20,
 col=rainbow(3))
```

... or discrete numeric data (supplied as a factor)...
```{r}
values <- c(0.01, 0.1, 0.2, 0.25, 0.5, 1)
xyloplot(
	replicate(n=5, simplify=FALSE, 
	expr=factor(sample(values, size=10, replace=TRUE), levels=values)), 
	col=rainbow(5))
```

... or general factors with wordy levels...
```{r}
xyloplot(
 sample(c("goldfish","cat","dog","fish","mouse","giraffe"), 
 size=100, replace=TRUE))
```

The `xylo_positions` function gives you the x-coordinates of the centres of the xylophones in case you want to add further graphical objects. For example, here we'll add the upper quartile as a red line for samples from 3 normal distributions.

```{r}
data <- lapply(1:3, function(mean) rnorm(mean=mean, n=1000))
xyloplot(x=data)
n <- length(data)
positions <- xylo_positions(n)
upper_qs <- sapply(data, function(sample) quantile(sample, prob=0.75))
segments(x0=positions-1/n/2, x1=positions+1/n/2, y0=upper_qs, y1=upper_qs, col="red", pch=19)
```

`xyloplot` can be used to plot population demographics-style histograms (i.e. 'split in the middle') using the `left_right_xylo` function. This function must be supplied a left-hand side vector (or list of vectors), `lhs`, and a right-hand side vector (or list of vectors of the same length as `lhs`), `rhs`. A length-2 vector of colours can then be used to colour the left and right-hand sides respectively.

```{r}
data_rhs <- lapply(2:4, function(mean) rnorm(mean=mean, sd=2, n=2000))
left_right_xylo(lhs=data, rhs=data_rhs, col=rainbow(2))
```

`xyloplot` accepts a `just` argument, a vector whose elements correspond to the elements of the list `x` (and which should therefore be possible to recycle to the same length as `x`), and whose components should take values in `{0, 0.5, 1}`. Values of `0` cause the bars of the resultant xyloplot to align to the pivots on the left (like a traditional histogram), values of `0.5` cause them to align centre align, and values of `1` cause them to align to the right. Note that graphical options, e.g. `col`, are recycled over the distributions going from left to right, regardless of whether they are aligned to the centre, left or right. 

`xyloplot` also accepts a `pivot` argument, which like `just` should be a vector whose elements correspond to the elements of `x`, and determine which pivot the histograms representing the distributions in each element of `x` should be assigned to.

Note that the `left_right_xylo` argument is thus a shorthand method of generating the equivalent plot with `xyloplot`...

```{r eval=FALSE}
xyloplot(c(data, data_rhs), pivot=rep(1:3, 2), just=rep(1:0, each=3), col=rep(rainbow(2), each=3))
```

Multiple xylophones can be placed on the same pivot by including duplicated elements in the `pivot` argument. Of course they would overlap each other, however the extent of the overlap can be seen if using non-opaque colours.
```{r}
xyloplot(list(rnorm(1000), rnorm(1000)+2), pivot=rep("a", 2), col=rainbow(n=2, alpha=0.5))
```

`xyloplot` accepts a `freq` argument. It is equivalent to the `freq` argument in `hist`: a logical value which if `FALSE` (default) makes the width of bars represent frequency densities, and if `TRUE`, makes the width of bars represent frequencies/counts. In the example below, note that the `rhs` object contains twice as many observations as the `lhs` object. Setting `freq` to `TRUE` will therefore mean that the right-hand side bars collectively occupy twice as much width as the left-hand side ones.

```{r echo=FALSE}
old_mfrow <- par("mfrow")
par(mfrow=c(1, 2))
```

```{r}
lhs <- rnorm(n=1000)
rhs <- rnorm(n=2000)
plot(main="freq=FALSE", left_right_xylo(lhs=lhs, rhs=rhs, col=rainbow(2), freq=FALSE))
plot(main="freq=TRUE", left_right_xylo(lhs=lhs, rhs=rhs, col=rainbow(2), freq=TRUE))
```

```{r echo=FALSE}
par(mfrow=old_mfrow)
```

## xyloplot objects

The `xyloplot` function returns an object of class `"xyloplot"`, containing information about the graphical elements required to construct the plot. Plots with default parameters are generated when the object is printed. In order to set some graphical parameters, the S3 generic `plot` function for xyloplots must be called on the object.

Examples include creating horizontal xyloplots (`vertical=FALSE`), removing the box around the plot which is drawn by default (`box=FALSE`), and setting the labels for the x and y axis. For more options, see the help page `?plot.xyloplot`.

```{r}
plot(
	main="Horizontal unboxed xyloplot", 
	vertical=FALSE, 
	box=FALSE, 
	pivots_lab="set this text with the `pivots_lab` parameter",
	value_lab="set this text with the `value_lab` parameter",
	x=xyloplot(x=rnorm(1000)))
```
 
