---
title: "Xyloplot examples"
author: "Daniel Greene"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Examples}
  %\VignetteEngine{knitr::rmarkdown}
  %\usepackage[utf8]{inputenc}
---
```{r echo=FALSE}
knitr::opts_chunk$set(dev="svg", fig.width=7, fig.height=5)
library(xyloplot)
set.seed(1)
```

The `xyloplot` function displays continuous or discrete data as histogram style violin plots, or "xylophones".

`xyloplot` accepts a simple numeric vector...
```{r}
xyloplot(rnorm(1000))
```

... or a list of simple numeric vectors...
```{r}
xyloplot(
 x=lapply(1:3, function(mean) rnorm(mean=mean, n=1000)), 
 breaks=20,
 col=rainbow(3), 
 main="title")
```

... or discrete numeric data (supplied as a factor)...
```{r}
xyloplot(
 replicate(n=5, simplify=FALSE, 
  expr=factor(sample(c(0.01, 0.1, 0.2, 0.25, 0.5, 1), size=10, replace=TRUE))), 
 col=rainbow(5))
```

... or general factors with wordy levels...
```{r}
xyloplot(
 sample(c("goldfish","cat","dog","fish","mouse","giraffe"), 
 size=100, replace=TRUE))
```

The `xylo_positions` function gives you the x-coordinates of the centres of the xylophones in case you want to add further graphical objects. For example, here we'll add the upper quartile as a red line for samples from 3 normal distributions.

```{r}
data <- lapply(1:3, function(mean) rnorm(mean=mean, n=1000))
xyloplot(x=data)
n <- length(data)
positions <- xylo_positions(n)
upper_qs <- sapply(data, function(sample) quantile(sample, prob=0.75))
segments(x0=positions-1/n/2, x1=positions+1/n/2, y0=upper_qs, y1=upper_qs, col="red", pch=19)
```
